Distributed Topology Discovery in Self-Assembled Nano Networks

### 2014-02-18 trunk/PAPER/implementation.tex
general fixing

A possible hardware implementation of the DiSR algorithm will be
described in this section.  Before enter in to detail of  DiSR
architecture,  an idea of node structure tailored for this particular
scenario, can be found depicted in Fig.~\ref{fig:node_structure}. In
particular, such node should have the following foundamental elements:

@@

\item I/O Buffers-Trasceivers: These elements, one for
each port,  are responsible of data transmission between neighbours
nodes. The packets received or sent are stored in specifieds buffers
named input and output buffers respectivelly.

\item Switch Matrix: driven by a switch controller, enables the
reciprocal connections beetween the devices inside the node. Before
the creation of segments, such controller receives information from
DiSR block.  After this phase, the switch controller will be driven by
the block wich implements the routing algorithm. The Switch matrix is
essentially a series of multiplexers and demultiplexers.

\item : Processing Element (PE):Is strictly related to the node
functionality  and role inside a given network, e.g. being a
computation or storage node.

\item DiSR-Routing: The DiSR block contains all the hardware, control
logic and configuration registers, related to the implementation of the
proposed approach. The routing algorithm receives information
from DiSR  wich indicates the status of the segments that interests a
particular node. The routing operation will take in account of these
informations in order to perform a deadlock-free routing. Both Routing
and DiSR are connected to the switch matrix in order to receive
packets from the input buffer. Since the packets are processed one at
a time, a specific arbiter should be present within the switch matrix
controller.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{pictures/node_structure.eps}
  \caption{A possible node structure tailored for a DNA nanonetwork.}
 \label{fig:node_structure}
\end{figure}
%--------------------------------------------------------------------
\subsection{DiSR Architecture}
\label{ssec:disr_architecture}

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/disr_rtl_updated.eps}
  \caption{\emph{DiSR} block architecture.}
 \label{fig:implementation}
\end{figure}

To give a basic estimate of the overhead needed we will focus on DiSR-
specific components, that is, the control logic and configuration
registers needed to implement DiSR. In Figure~\ref{fig:implementation}
is shown a sketch of the  architecture, which mainly consist in the
following building elements:

@@

	\item \textbf{DBS block:} 
	It takes trace of the DBS state machine, 
	consisting of a 3-bit register (to cover six DBS values) and the 
	required combinational logic. This block receives signals from 
	stored packets  in order to decode the packet type, and from 
	the control circuitry to change its status.
	The bootstrap node is selected by setting the boot signal to high
	during the initialization phase.

	\item \textbf{LED registers:} 
	A set of registers store the \emph{LED} listed in the
	section~\ref{sec:disr_concepts}.  In particular, in the
	link\_visited[] table, can be pointed out the presence of two  
	particular registernamed \emph{tflag} and \emph{busy}.  The Latter
	is used to indicate if a specific port is alredy visited or tvisited and
	the former indicates specifically if the node is visited or tvisited.

    \item \textbf{Control circuitry:}
    This circuitry decodes the the      incoming packet, the LED
	registers and the DBS,     updating them when required. This block
	drives the LED register   to store or, to read the LED data
	according with     \emph{write enable (WE)} or \emph{read enable (RE)}
	signals. Then, the resulting \emph{Ctrl-Out} drives the other
	communication resources for actuating the DiSR routing operation.

	\item \textbf{Packager:}
	Essentially is a set of registers and multiplexers. Starting from
	an incoming packet stored in the input buffer, it updates the
	content of the packet before the retransmission to   a destination
	node.  In some cases, this circuitry build the packet from the
	scratch.  For example this happens when a node is bootsrap and for
	the first time it should inject a
	\emph{strating\_segment\_request}. In Fig.~\ref{fig:packager} is
	shown the architecture of such packager block. When a packet is
	created for the first time, the multiplexers driven by the
	controller select the information coming from the internal node
	such as the \emph{soruce\_ID}. In this situation the TTL field
	should be resetted to its initial value and the \emph{segment\_ID}
	is composed by the node ID and by the output port identification 
	number.

@@

While the elements mentioned above are part of the DiSR circuitry, 
the other devices depicted in Fig.~\ref{fig:implementation} like the 
multiplexer and demultiplexer are implemented outside DiSR.
In particular, such components will be within the switch matrix 
(refers to the Fig~\ref{fig:node_structure}).

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/packager.eps}
  \caption{Packager block}
 \label{fig:packager}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/pipeline.eps}
  \caption{Pipeline description of DiSR operations}
 \label{fig:pipeline}
\end{figure}

Each just described building block, reacts with the rising edge  of
the\emph{clock signal}. An asynchronous system \emph{reset} signal is
also present for restoring all registers to their default values. When
this last signal is set, the DBS state machines for each network's
node are set as \emph{Free}.  
In Fig.~\ref{fig:pipeline} is described the pipeline structure of the
DiSR architecture. The packet incoming from the neighbours node, or
locally generated will be send to the DBS state machine in order to te
ensure the status commutation. While the DBS update its state,  in the
next clock cycle the control circuitry will update the Led status and
will drive the sitch matrix in order to send eventually the packets to
the  right output buffer.
%--------------------------------------------------------------------
\subsection{Synthesis results}

As discussed above, one of the main design challenge of DNA Self-
Assembled systems is the limited resource available to implement both
computations and routing decisions for each node. Assuming a budget of
$10^4$ CNFETs for each network's node~\cite{liu_jetcs}  we estimated
the required resources for implementing the entire DiSR block. The RTL
description of the circuitry described in the last section (with an
hardware description language)  of the DiSR circuitry has been written
and synthesized at gate-level. Considering the specific layout of each
single logic elements (NAND, full-adder, latch etc.), it has been
possible to get a rough estimate of the number of transistors
necessary to implement the DiSR logic. Figure~\ref{fig:imp_trend}
shows the results of synthesis in terms of number of devices (CNFETs)
versus the number of network nodes while the network scales up from
$10\times10$ to $100\times100$ nodes.

\begin{figure}
  \centering
  \includegraphics[width=0.50\textwidth]{pictures/synthesis.eps}
  \caption{Number of network's node (N) vs number of CNFETs.}
 \label{fig:synthesys}
\end{figure}

In particular, observing the Figure~\ref{fig:imp_trend} can 
be pointed out that the proposed implementation occupies about 17,5\% of the node budget. In particular the main contribution is due to the controller circuitry followed by LED registers.
Further, more than the absolute number of devices itself, it is interesting to
observe that the circuitry
complexity which implements the DiSR algorithm increases with a slowly
growing trend.  
An intuitive explanation of this is the relatively simple logic of
DiSR, whose behaviour is almost code on scalable storage structures.
For example, the number of registers
implementing the \emph{link\_visited[]} and \emph{link\_tvisited[]}
table follow the logarithmic function $N_{reg}=N_{port} \cdot
log_2(N)$ where \emph{Nport} is the number of the router’s ports, N is
the number of network’s nodes.

In order ha have estimate the maximum working clock frequency for the implemented
circuitry, the \cite{deng_isscc07}

>>>>

A possible hardware implementation of \disr{} algorithm will be
described in this section. Figure~\ref{fig:node_structure} depicts the general
structure of a node in the particular scenario of DNA nano Network-On-Chip. In particular such a node
is composed by the following fundamental elements:

@@

	\item \textbf{I/O Buffers-Transceivers:} These elements, one for
	each port,  are responsible for data transmission with neighbors
	nodes. The packets received or sent are stored in specific buffers
	named input and output buffers respectively.

	\item \textbf{Switch Matrix:} Driven by a switch controller, it enables the
	reciprocal connections among devices inside the node. Before segments'
	creation, this controller receives information from
	\disr{} block. After this phase, the switch controller will be driven by
	the block implementing the routing algorithm. The \emph{Switch Matrix} is
	essentially composed of a series of multiplexers and demultiplexers.

	\item  \textbf{Processing Element (PE):} Is strictly related to the
	functionality and role that the node cover inside a given network (e.g. being a
	computation or storage node).

	\item \textbf{\disr{}-Routing:} The \disr{} block contains all the hardware, control
	logic and configuration registers, needed by the implementation of the
	proposed approach. The routing algorithm receives information
	from \disr{}  which indicates the status of segments related to a
	particular node. Routing operations will take into account these
	information to obtain deadlock freedom. Both Routing
	and \disr{} are connected to the \emph{Switch Matrix} in order to receive
	packets from the input buffer. Since the packets are processed one at
	a time, a specific arbiter should be present within the \emph{Switch Matrix}
	controller.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{pictures/node_structure.eps}
  \caption{A possible node structure tailored for a DNA nano Network-On-Chip.}
 \label{fig:node_structure}
\end{figure}
%--------------------------------------------------------------------
\subsection{\disr{} Architecture}
\label{ssec:disr_architecture}

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/disr_rtl_updated.eps}
  \caption{\emph{\disr{}} block architecture.}
 \label{fig:implementation}
\end{figure}

To give a basic estimate of the overhead needed we focus on \disr{}-
specific components namely the control logic and configuration
registers needed to implement \disr{}. Figure~\ref{fig:implementation}
shows an architecture sketch, which mainly consist in the
following building elements:

@@

	\item \textbf{DBS block:}

	It takes trace of the DBS state machine, consisting of a 3-bit
	register (to cover six DBS values) and the  required combinational
	logic. This block receives signals from  stored packets  in order
	to decode the packet type, and from  the control circuitry to
	change its status. The bootstrap node is selected by setting the
	boot signal to high during the initialization phase.

	\item \textbf{LED registers:}
	A set of registers store the \emph{LED} defined in
	section~\ref{sec:disr_concepts}.  In the
	link\_visited[] table, we defined two specific registers named
	\emph{tflag} and \emph{busy}. While the first one indicates
    if the specific port of a segment is already candidate or assigned,
    the second one records if a specific port is already \emph{visited} or \emph{tvisited}.

    \item \textbf{Control circuitry:}
    This circuitry decodes incoming packet, LED
	registers and DBS, updating them when required. This block
	drives LED register according with \emph{write enable (WE)} or \emph{read enable (RE)}
	signals. Then, the resulting \emph{Ctrl-Out} drives the other
	communication resources to actuate \disr{} routing operations.

	\item \textbf{Packager:}
	Essentially it is a set of registers and multiplexers. Starting from
	an incoming packet stored in the input buffer, it updates packet's
	content before retransmission to a destination
	node.  In some cases, this circuitry builds the packet from
	scratch e.g. when a node is bootstrap and for
	the first time it should inject a \emph{starting\_segment\_request}. 
	Figure~\ref{fig:packager} depicts the architecture of this
	block. When a packet is created for the first time, the multiplexers 
	driven by the controller, will select information incoming from an internal 
    node such as the \emph{source\_ID}. In this situation the 
	TTL field should be reset to its initial value and the \emph{segment\_ID}
	is composed by the node ID and by the output port identification 
	number.

@@

While the elements mentioned above are part of the \disr{} circuitry,
the other devices depicted in Figure~\ref{fig:implementation} like the
multiplexer and demultiplexer are implemented outside \disr{}. In
particular, such components will be inside the \emph{Switch Matrix} (see Figure~\ref{fig:node_structure}).

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/packager.eps}
  \caption{Packager block.}
 \label{fig:packager}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.40\textwidth]{pictures/pipeline.eps}
  \caption{Pipeline description of \disr{} operations.}
 \label{fig:pipeline}
\end{figure}

All the blocks described above, reacts with the rising edge of
the \emph{clock signal}. An asynchronous system \emph{reset} signal is
also present to restore all registers to their default values. When
this signal is set, DBS state machines for each network
node are set as \emph{Free}.  
Figure~\ref{fig:pipeline} is shows the pipeline structure of the
\disr{} architecture. Packets incoming from neighbor nodes (or
generated locally) are sent to the DBS state machine in order to
ensure the status commutation. While DBS updates its state, in the
next clock cycle the control circuitry will update LED registers status and
will drive the \emph{Switch Matrix} in order to send eventually the packets to
the right output buffer.
%--------------------------------------------------------------------
\subsection{Synthesis results}

As discussed above, one of the main design challenges of DNA Self-
Assembled systems is the limited number of resources available to implement both
computations and routing decisions for each node. Assuming a budget of
$10^4$ CNFETs for each network's node~\cite{liu_jetcs}  we estimated
the required resources to implement the entire \disr{} block. The RTL
description of the circuitry described in the last section (with an
hardware description language)  of the \disr{} circuitry has been written
and synthesized at gate-level using \emph{Synopsys Design Compiler} with 
a generic technology library (GTECH).
Considering the specific layout of each single logic elements  (NAND,
full-adder, latch etc.), it has been possible to get a rough estimate
of the number of transistors necessary to implement \disr{} logic.
Figure~\ref{fig:synthesys} shows the results of synthesis in terms of
number of devices (CNFETs) versus the number of network nodes while
the network scales up from $10\times10$ to $100\times100$ nodes.

\begin{figure}
  \centering
  \includegraphics[width=0.50\textwidth]{pictures/synthesis.eps}
  \caption{Network size vs number of CNFETs necessary to implement
  the \disr{} circuitry.}
 \label{fig:synthesys}
\end{figure}

In particular, Figure~\ref{fig:synthesys} shows that the proposed 
implementation occupies about 17,5\% of our node budget. 
The main contribution is due to the controller circuitry followed by LED registers. 
Further, more than the
absolute number of devices itself, it is interesting to observe that
the circuitry complexity which implements \disr{} algorithm increases
with a slowly growing trend. An intuitive explanation of this is the
relatively simple logic of \disr{}, whose behavior is almost coded on
scalable storage structures. For example, the number of registers
implementing the \emph{link\_visited[]} and \emph{link\_tvisited[]}
table follow the logarithmic function $N_{reg}=N_{port} \cdot
log_2(N)$ where \emph{Nport} is the number of the router’s ports and N is
the number of network’s nodes.

In order to have an idea of the maximum working clock frequency for
the implemented circuitry, timing results can be obtained from a 
gate-level netlist. Since a complete technology library useful for
commercial synthesis tool (e.g. Design Compiler) is not available for
this particular technology, we have
synthesized the RTL description of \disr{}, with a standard 32~nm CMOS
library, obtaining a delay of about $\mathrm{\tau_{\disr{}}=10} \ FO4$
(fan-out of four). Considering the results obtained
in~\cite{deng_isscc07}, which reports the ratio in terms of FO4
between a standard 32~nm CMOS technology and a carbon nanotube ones, 
rough delay estimation can be obtained. In particular in such work 
has been reported a ratio of:  
\begin{equation}
\mathrm{\frac{FO4_{CMOS}}{FO4_{CNFET}}=2}  
\end{equation}
Considering that F04 CMOS is about 30~ps, and the results of last
equation, a FO4 for the CNFET case is equal to about 15~ps, the working 
frequency can be computed as:
\begin{equation}
\mathrm{f_{clk}=\frac{1}{T_{max}}=\frac{1}{\tau_{\disr{}} \cdot FO4_{CNFET}}}=
\frac{1}{10\cdot 15 \cdot 10^{-12}}=6,6 GHz
\end{equation}
where $\mathrm{\tau_{\disr{}}}$ is expressed in terms of FO4.

Regarding the power consumption of whole \disr{} devices, 
we avoided an accurate power analysis with the switch activity
annotation due to the fact that \disr{} circuitry will be active 
only once during system startup. Indeed, after the setup phase, 
when all segments are mapped, the \disr{} block will stop its activity 
passing all the information related to segments to the routing
algorithm. For this reason the dynamic power falls to zero and
the only consumption could be due to the leakage
consumption.

### 2014-02-18 trunk/PAPER/simulation.tex
general fixing

In this section we will test the proposed \disr{} approach to demonstrate
its effectiveness and compare it against topology agnostic approach
based on spanning trees and
broadcasting~\cite{Patwardhan05evaluatingthe}, to measure how \disr{}
performs in covering the network structure.

@@

In order to quantitatively and qualitatively evaluate the proposed approach a
specific simulation environment has been developed, resulting in
the open source and freely available project called
Nanoxim~\cite{nanoxim}.
Nanoxim is a SystemC tool based on a almost rewritten
version of the Noxim Network-on-Chip simulator~\cite{noxim}. While some
complex features have been removed (e.g. wormhole, congestion/topology
aware routing and selection strategies) new features specifically
tailored for the nanoscale scenario were introduced, e.g. the ability to simulate a random
network, the implementation of the \disr{} to obtain the segment topology
and the support for defective links and nodes.

%------------------------------------------------------------------------------
\subsection{Experimental setup}
The following parameters have been taken into account when
performing the \disr{} simulation:
\begin{itemize}
\item {\emph{Size of the network}}: number of nodes, on a range from
10x10 to 100x100 sized networks. 
%\item {\% defective links}: the probability that each link is
%disconnected or not present, from 0 the amount that yields
%no more segments.
\item {\emph{\% defective nodes}}: the probability that a node is not
working, thus having all its links cannot be utilized during \disr{} setup.  
%Same range as defective links parameter.
\item {\emph{Bootstrap node}}: the node used from upper layer 
to inject the \disr{} process. When not explicitly investigating the
impact of each single bootstrap choice, a set representative regions have been
considered, e.g. the central part of the network and the edge corners.
\end{itemize}
To present the results, the following evaluation metrics have been adopted:
\begin{itemize}
\item{Node coverage}: this is the fraction of nodes that
are assigned to a segment. In the ideal case, all the non defective
nodes should be assigned, so this metric is useful to show how some
disconnected regions can negatively impact on the whole \disr{}
effectiveness.
\item{\emph{Latency}}: this measures how the cycles required to complete the
segment assignment scales for increasing network sizes and defect rates
\item{\emph{Bootstrap node effect}}: this evaluates the impact of the chosen
bootstrap node on the node coverage.
\end{itemize}

@@

In this section we analyze the results in terms of node
coverage and latency at different network sizes, defect rates and
bootstrap injection points.  In particular
Figure~\ref{fig:results_coverage} shows
node coverage for \disr{} and RPF tree based approach respectively. While
the first aim of \disr{} is not to reach the optimal coverage, we still
can observe a quite good performance as compared to the tree based
approach. Note that defect rates beyond 25\% lead to many disconnected
regions of nodes that \disr{} currently cannot handle.
For example Figure~\ref{fig:net} show a 30x30 network with a 25\%
defect in which the bottom-left part cannot be reached due
disconnected regions. Note that remaining defective nodes belonging to
connected regions are successfully surrounded by \disr{} coverage; in any
case, these defect levels should be considered as worst case scenarios, so the
achieved coverage of $0.5$ is a satisfying result for this first
version of \disr{}. On the other hand, the network size seems to have a
limited impact when defect rate do not introduce too much disconnected
regions. 
\begin{figure*}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{pictures/set1.eps} & 
\includegraphics[width=0.48\textwidth]{pictures/coverage.eps} \\
(a) & (b) 
\end{tabular}
\caption{\disr{} (a) and RPF (b) node coverage}
\label{fig:results_coverage}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.70\textwidth]{pictures/net.ps}
\caption{Covered regions in a 30x30 network with 25\% defects}
\label{fig:net}
\end{figure*}

@@

The number of cycles required to complete segment mapping process
is shown in Figure~\ref{fig:results_latency}. In this case comparison
against tree-based shows better (lower) values at different defect
rates.  Rather than the absolute numbers, what it's more interesting
to observe is how \disr{} latency scales with network size. For example,
going from 900 to 2500 nodes, at the medium defect rate of $0.15$,
leads to an increase from 3000 to 4500 cycles. It should be noticed
also how the effect of defect rate is increasing until the threshold
of $0.25$ is reached, meaning that until that limit \disr{} finds it more
and more difficult to complete the process due increasing defective
paths, but still discover new segments when let running for a more
extended amount of cycles. This behavior is not reported in the RPF
based approach, meaning that a tree based approach, although starting
from higher values, is less affected by defect rates (when low rates
are considered).
After the $0.25$ threshold, the impact of entire disconnected regions
becomes predominant and both approaches become faster in completing
the covering process, since far less nodes can be actually reached.

Finally, Figure~\ref{fig:results_bootstrap} visually represents  the
stability of the approach against choices of a different bootstrap
node in a 10x10 network.  This is an important aspect to evaluate when
considering that one of the main advantages of \disr{} against all the
tree based approaches should be the ability of choosing whatever
bootstrap node, without having to care about the role assumed in the
future by the chosen root node.  In other words, after segments have
been established, the bootstrap node is like every other node, i.e. it
is not center of a structure, and it is not an hotspot for the traffic
distribution. The results in terms of coverage,  shown for low, medium
and high defect profiles, demonstrate a relatively limited
impact of the bootstrap choice in the low/medium scenarios, while a
$30\%$ instability is found for very high defect rates. This also
sounds acceptable, since when a lot of defective nodes are present, the
particular position of the bootstrap node could lead to a completely
different evolution in the \disr{} setup process.

@@

% TODO: long paper
\subsection{Other Optimizations}
Some optimization parameters, which demonstrated to improve the \disr{}
results have been fixed to some good values, but are not subject to
further investigation; once again the focus here is not the
optimal set of segments, but just demonstrating a working approach. 
These are optimizations are:
\begin{itemize}
\item \emph{cycle\_links}: max number of retries across the set of links of
each node. While searching for a free link due an incoming request,
the same request is cancelled after a given number of tries. This
gives to preceding node change to test a different paths. Default is set to 1.
\item \emph{bootstrap\_timeout}: number of time units that a bootstrap node
should wait before assuming that a livelock in the starting segment
process has occurred. In the worst case, we can imagine that longest
path required is the one returning to the bootstrap node after having
traveled across all the links. So, although this is just an extreme
situation, a good upper limit can be safely be set to NxN.
\item \emph{bootstrap\_immunity}: in order to avoid the failure of the whole \disr{}
setup process, a bootstrap node should not have defective links.
Enabling this optimization, a bootstrap node is immune to defects.
We may think of a pre-bootstrap phase that properly selects (from upper
layer via) a bootstrap node which is tested as properly connected. We
enabled this optimization, however empirical tests showed us that only
simulations using bootstrap nodes placed on edges would be heavily
affected, since these are nodes starting with a lower number of links,
e.g. corner nodes could only have two connected direction, so even a
defective link could prevent starting segment packet to return to
bootstrap to close the loop a create the segment.
\end{itemize}

>>>>

In this section we will test the proposed \disr{} approach to demonstrate
its effectiveness and compare it against a topology agnostic approach
based on spanning trees and
broadcasting~\cite{Patwardhan05evaluatingthe}, to measure how \disr{}
performs in covering the network structure.

@@

In order to quantitatively and qualitatively evaluate the proposed approach a
specific simulation environment has been developed, resulting in
the open-source and freely available project called
Nanoxim~\cite{nanoxim}.
Nanoxim is a SystemC tool based on an almost rewritten
version of the Noxim Network-on-Chip simulator~\cite{noxim}. While some
complex features have been removed (e.g. wormhole, congestion/topology
aware routing and selection strategies) new features specifically
tailored for the nanoscale scenario were introduced, e.g. the ability to simulate a random
network, the implementation of \disr{} to obtain the segment topology
and the support for defective links and nodes.

%------------------------------------------------------------------------------
\subsection{Experimental setup}
The following parameters have been taken into account when
performing the \disr{} simulation:
\begin{itemize}
\item {\emph{Size of the network}}: number of nodes, on a range from
10x10 to 100x100 sized networks. 
%\item {\% defective links}: the probability that each link is
%disconnected or not present, from 0 the amount that yields
%no more segments.
\item {\emph{\% defective nodes}}: the probability that a node is not
working, thus having all its links not able to be utilized during \disr{} setup.  
%Same range as defective links parameter.
\item {\emph{Bootstrap node}}: the node from upper layer that
injects the \disr{} process. When not explicitly investigating the
impact of each single bootstrap choice, a set representative regions have been
considered, e.g. the central part of the network and the edge corners.
\end{itemize}
To present the results, the following evaluation metrics have been adopted:
\begin{itemize}
\item{\emph{Node coverage}}: this is the fraction of nodes that
are assigned to a segment. In the ideal case, all the non defective
nodes should be assigned, so this metric is useful to show how some
disconnected regions can negatively impact on the whole \disr{}
effectiveness.
\item{\emph{Latency}}: this measures how the cycles required to complete the
segment assignment scale for increasing network sizes and defect rates.
\item{\emph{Bootstrap node effect}}: this evaluates the impact of the chosen
bootstrap node on the node coverage.
\end{itemize}

@@

In this section we analyze the results in terms of node
coverage and latency with different network sizes, defect rates and
bootstrap injection points. In particular
Figure~\ref{fig:results_coverage} shows
node coverage for \disr{} and RPF tree based approach respectively. While
the first aim of \disr{} is not to reach the optimal coverage, we still
can observe a quite good performance as compared to the tree based
approach. Note that defect rates beyond 25\% lead to many disconnected
regions of nodes that \disr{} currently cannot handle.
For example Figure~\ref{fig:net} show a 30x30 network with a 25\%
defect in which the bottom-left part cannot be reached due to
disconnected regions. Note that the remaining defective nodes belonging to
connected regions are successfully surrounded by \disr{} coverage; in any
case, these defect levels should be considered as worst case scenarios, so the
achieved coverage of $0.5$ is a satisfying result for this first
version of \disr{}. On the other hand, the network size seems to have a
limited impact when defect rate does not introduce too many disconnected
regions. 
\begin{figure*}
\centering
\begin{tabular}{cc}
\includegraphics[width=0.48\textwidth]{pictures/set1.eps} & 
\includegraphics[width=0.48\textwidth]{pictures/coverage.eps} \\
(a) & (b) 
\end{tabular}
\caption{\disr{} (a) and RPF (b) node coverage}
\label{fig:results_coverage}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.70\textwidth]{pictures/net.ps}
\caption{Covered regions in a 30x30 network with 25\% of defects}
\label{fig:net}
\end{figure*}

@@

The number of cycles required to complete segment mapping process
is shown in Figure~\ref{fig:results_latency}. In this case the comparison
against the tree-based approach shows better (lower) values at different defect
rates. Rather than the absolute numbers, what is more interesting
to observe is how \disr{} latency scales with network size. For example,
going from 900 to 2500 nodes, at the medium defect rate of $0.15$,
leads to an increase from 3000 to 4500 cycles. It should be noticed
also how the effect of defect rate is increasing until the threshold
of $0.25$ is reached, meaning that until that limit \disr{} finds 
the completion of the process more and more difficult due to the increasing defective
paths, but it still discover new segments if it runs for an
extended amount of cycles. This behavior is not reported in the RPF
based approach, meaning that a tree-based approach, although starting
from higher values, is less affected by defect rates (when low rates
are considered).
After the $0.25$ threshold, the impact of disconnected regions
becomes predominant and both approaches become faster in completing
the covering process, since far less nodes can be actually reached.

Finally, Figure~\ref{fig:results_bootstrap} visually represents the
stability of the approach against a different bootstrap node choice
in a 10x10 network. This is an important aspect to evaluate 
considering that one of the main advantages of \disr{} against all the
tree-based approaches is the possibility of choosing whatever
bootstrap node, without having to care about the role assumed in the
future by the chosen root node. In other words, after the segments have
been established, the bootstrap node is like every other node, i.e. it
is not center of a structure, and it is not an hotspot for the traffic
distribution. The results in terms of coverage, shown for low, medium
and high defect profiles, demonstrate a relatively limited
impact of the bootstrap node choice in the low/medium scenarios, while a
$30\%$ instability is found for very high defect rates. This also
sounds acceptable, since when a lot of defective nodes are present, the
particular position of the bootstrap node could lead to a completely
different evolution in the \disr{} setup process.

@@

% TODO: long paper
\subsection{Other Optimizations}
Some optimization parameters, which demonstrated to improve the \disr{}
results, have been fixed to some good values and are not subject to
further investigation in this paper; once again the focus here is not the
optimal setup of segments, but just demonstrating a working approach. 
These parameters are:
\begin{itemize}
\item \emph{cycle\_links}: max number of retries across the set of links of
each node. While searching for a free link due to an incoming \texttt{SEGMENT\_REQUEST},
the request itself is cancelled after a given number of tries. \hl{This
gives to the preceding node, that receives a \texttt{SEGMENT\_CANCEL}, the
chance to test a different path instead of waiting indefinitely.}
Default value is set to 1, i.e. each link is tried one time.
\item \emph{time to live (TTL)}: when a segment request is cancelled
along a certain path (using a \texttt{SEGMENT\_CANCEL}), a TTL field is
decreased in the request packet, so that requests that have been
denied too much times can desist, even if free links are
available. This is not only an optimization to shorten
requests' processes, but also solves some critical situations in which
request packets are blocked in routing loops in a particular path.
\item \emph{bootstrap\_timeout}: number of time units that a bootstrap node
should wait before assuming that a livelock in the starting segment
process has occurred. In the worst case, we can imagine that longest
path required is the one returning to the bootstrap node after having
traveled across all the links. So, although this is just an extreme
situation, a good upper limit can be safely be set to $N \times N$.
\item \emph{bootstrap\_immunity}: in order to avoid the failure of the whole \disr{}
setup process, a bootstrap node should not have defective links.
Enabling this optimization, a bootstrap node is immune to defects.
We may think about a pre-bootstrap phase that properly selects (from upper
layer via) a bootstrap node which is tested as properly connected. We
enabled this optimization, however empirical tests have shown us that only
simulations using bootstrap nodes placed on edges would be heavily
affected by similar issues since these nodes start with a lower number of links,
e.g. corner nodes could only have two connected direction, so even a \hl{single}
defective link could prevent a starting segment packet from coming back to the
bootstrap node to close the loop and create the \hl{first} segment.
\end{itemize}
